Extended Formal Analysis of the EDHOC Protocol in Tamarin

### 2022-07-23 formalization.tex

\begin{figure*}[ht]
\begin{align*}
    \mPredPcs \triangleq\ & \forall \mTID, I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, 
\mRComplete^{t_2}(\mTID, 
I, R, 
\mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
	&\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land 
\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, I, R) 
\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
%    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) 
%\land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
%
    \mPredInjI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, 
S) \land t_1 \lessdot t_2
    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', 
I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))\\[2em]
%
    \mPredInjR \triangleq\ &
    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot 
\mRComplete^{t_2}(\mTID_R, 
I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, 
S) \land t_1 \lessdot t_2
    \land (\forall \mTID_R', I' R' t_1' \mLogicDot 
\mRComplete^{t_1'}(\mTID_R', 
I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, 
\mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \land t_1 \lessdot 
t_2))\\[2em]
%
    \mPredImpI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot 
\mIComplete^{t_1}(\mTID_I, 
I, R, \mSessKey, 
S)
    \rightarrow\\
      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, 
I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S')\\
      &\ \ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq 
t_1
        ))\\
    &\lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land 
\mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \land t_1 \lessdot t_2))
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

%----------------------------------------------------------------------- PFS
\subsubsection{Secrecy of the Session-Key Material.}
\label{sec:secrecy}
We show that the adversary cannot gain access to the session-key material
\mSessKey{}, even under the weak post-compromise security model which allows 
the
adversary limited access to the long-term keys of the involved parties via
a
secure interface before the session starts, and unrestricted access to the long-term
keys of all parties as well as access to all other session keys after the session ends.
%
This property is formalized as \mPredPcs{} in Figure~\ref{fig:props}.
%

In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' which uniquely identifies the
current session, $I$ represents the identity of the initiator,
and $R$ represents the identity of the party who the initiator believes is
playing the responder role, while \mSessKey{} stands for the established
session-key material.
%
\mRComplete{} has analogous parameters; in particular, the responder $R$
believes $I$ is the party playing the initiator role.
%
Intuitively, the PCS property states that an adversary obtains \mSessKey{}
only if one of the following conditions hold: either a party's long-term key
is compromised before their run ends, or \mSessKey{} is directly revealed to
the attacker, or the TEE interface for the responder is used during the current
session for the initiator, i.e., after the initiator starts and before it
completes, or, similarly, the TEE interface for the initiator is used during the
current session for the responder.
%
Note that the properties in Figure~\ref{fig:props} are unlike the actual \mTamarin{} lemma in one minor way: \mTamarin's logic does not allow disjunctions to appear on
the left-hand side of an implication inside a universally-quantified formula.
%
Therefore, in the \mTamarin{} code, instead of using the disjunction
$\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\,
\mRComplete^{t_2}(\mTID, I, R, \mSessKey)$
to model the fact that either party may have completed their execution, we use
a single action parametrized by the terms $\mTID$, $I$, $R$, and \mSessKey.

@@

\begin{scriptsize}
\begin{verbatim}
1 lemma secrecyPFS:
2    all-traces
3    "All u v sk tid #t3 #t2.
4       (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
5       ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
6       | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
7       | (Ex #t1. EphKeyRev(sk)@t1)
8       | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
9       | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
10    )"
\end{verbatim}
\end{scriptsize}
%

In this formalization, we use the action \mT{CompletedRun(u, v, sk)}
(in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$
and $\mRComplete^{t_{2}}$.
%
As expected, this action is emitted by both \mT{I3} and \mT{R4}.
%
Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of
the ephemeral key for $I$ or $R$ or both.
%
Lines 8 and 9 captures that the parties TEEs must be inaccessible to the
adversary between the start and end of the execution as seen by each party
respectively.
%

>>>>

\begin{figure*}[ht]
\begin{align*}
    \mPredPcs \triangleq\ & \forall \mTID, I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, 
\mRComplete^{t_2}(\mTID, 
I, R, 
\mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
%
    \mPredInjI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, 
S) \land t_1 \lessdot t_2
    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', 
I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)\\[2em]
%
    \mPredInjR \triangleq\ &
    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot 
\mRComplete^{t_2}(\mTID_R, 
I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, 
S) \land t_1 \lessdot t_2
    \land (\forall \mTID_R', I' R' t_1' \mLogicDot 
\mRComplete^{t_1'}(\mTID_R', 
I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land t_0 \lessdot t_1)\\[2em]
%
    \mPredImpI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot 
\mIComplete^{t_1}(\mTID_I, 
I, R, \mSessKey, 
S)
    \rightarrow\\
      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, 
I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S')\\
      &\ \ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq 
t_1
        ))\\
    &\lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land t_0 \lessdot t_1)
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

%----------------------------------------------------------------------- PCS
\subsubsection{Secrecy of the session key material.}
\label{sec:secrecy}
We show that the adversary cannot gain access to the session-key material
\mSessKey{}, even under the weak post-compromise security model which allows 
the
adversary limited access to the long-term keys of the involved parties via
a
secure interface before the session starts, and unrestricted access to the long-term
keys of all parties as well as access to all other session keys after the session ends.
%
This property is formalized as \mPredPcs{} in Figure~\ref{fig:props}.
%

In the \mIComplete{} event, \mTID{} represents a ``thread identifier'' for the
current session and is used to correlate multiple events for the same run of the 
protocol for each party, $I$ represents the identity of the initiator,
and $R$ represents the identity of the party who the initiator believes is
playing the responder role, while \mSessKey{} stands for the established
session-key material.
%
\mRComplete{} has analogous parameters; in particular, the responder $R$
believes $I$ is the party playing the initiator role.
%
Intuitively, the PCS property states that an adversary obtains \mSessKey{}
only if one of the following conditions hold: either a party's long-term key
is compromised before their run ends, or \mSessKey{} is directly revealed to
the attacker, or the TEE interface for the responder is used during the current
session for the initiator, i.e., after the initiator starts, or, similarly, the 
TEE interface for the initiator is used during the
current session for the responder.
%
We use the thread identifier \mTID for each party to relate the starting event
to the corresponding commit event.
%
The property in Figure~\ref{fig:props} is unlike the actual \mTamarin{} lemma
in one minor way: \mTamarin's logic does not allow disjunctions to appear on
the left-hand side of an implication inside a universally-quantified formula.
%
Therefore, in the \mTamarin{} code, instead of using the disjunction
$\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\,
\mRComplete^{t_2}(\mTID, I, R, \mSessKey)$
to model the fact that either party may have completed their execution, we use
a single action parametrized by the terms $\mTID$, $I$, $R$, and \mSessKey.

@@

\begin{scriptsize}
\begin{verbatim}
1 lemma secrecyPCS:
2    all-traces
3    "All u v sk tid #t3 #t2.
4       (K(sk)@t3 & CompletedRun(u, v, sk, tid)@t2) ==>
5       ( (Ex #t1. LTKRev(u)@t1 & #t1 < #t2)
6       | (Ex #t1. LTKRev(v)@t1 & #t1 < #t2)
7       | (Ex #t1. EphKeyRev(sk)@t1)
8       | (Ex m1 #t0 #k. I1(tid, u, v, m1)@t0 & TEE(v)@k & (t0 < k | k < t2))
9       | (Ex m1 m2 #t0 #k. R2(tid, v, m1, m2)@t0 & TEE(u)@k & (t0 < k | k < t2))
10    )"
\end{verbatim}
\end{scriptsize}
%

In this formalization, we use the action \mT{CompletedRun(u, v, sk)}
(in line 4) to represent the disjunction of the events $\mIComplete^{t_{2}}$
and $\mRComplete^{t_{2}}$.
%
As expected, this action is emitted by both \mT{I3} and \mT{R4}.
%
Similarly, the action \mT{EphKeyRev(sk)} in line 7 stands for the reveal of
the ephemeral key for $I$ or $R$ or both.
%
Lines 8 and 9 captures that the parties TEEs must be inaccessible to the
adversary after the start of the session as seen by each party
respectively.
%
